---
title: "SICB"
author: "Nina Hernandez"
date: "2022-12-06"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Stat2Data)
library(glmmADMB)
library(utils)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(pscl)
library(MASS)
library(performance)
library(devtools)
library(R2admb)
```


```{r}
# exporting csv experiment 2 data
exp2 <- read.csv("exp2.csv")

# omitted all N/A values 
na.omit(exp2)
```

```{r}
#View(exp2)
```


# Making a model for the total major movements behavior 

```{r}
#selected for only treatment and movement columns. short for "movements data set"
mds <- exp2 %>%
  select(treatment, total_major_movements, hour, terrarium, video_name)
```

```{r}
# made terrarium variable a factor
mds <- transform(mds,fterrarium=factor(terrarium))
```

```{r}
# using the glmmadmb package to create a zero inflation mixed linear binomial plot 
mds = glmmadmb(formula = total_major_movements ~ treatment * hour + (1 | fterrarium), data = mds, family = "nbinom", zeroInflation = TRUE) 
summary(mds)
```
```{r}
# verifying model
hist(exp2$total_major_movements)
```
```{r}
poisson.test(sum(exp2$total_major_movements), length(exp2$total_major_movements))
```



# Making a model for the touches behavior 

```{r}
#selected for only the touches behavior. short for "touches data set"
tds <- exp2 %>%
  select(treatment, touches, hour, terrarium, video_name)
```

```{r}
# made terrarium variable a factor
tds <- transform(tds,fterrarium=factor(terrarium))
```

```{r}
# using the glmmadmb package to create a zero inflation mixed linear poisson plot 
tds = glmmadmb(formula = touches ~ treatment * hour + (1 | fterrarium), data = tds, family = "poisson", zeroInflation = TRUE) 
summary(tds)
```

```{r}
# verifying model
hist(exp2$touches)
```

```{r}
poisson.test(sum(exp2$touches), length(exp2$touches))
```


# Making a model for the foot-flagging behavior 

```{r}
#selected for only the foot flagging behavior. short for "foot-flags data set"
footds <- exp2 %>%
  select(treatment, foot_flags, hour, terrarium, video_name)
```

```{r}
# made terrarium variable a factor
footds <- transform(footds,fterrarium=factor(terrarium))
```


```{r}
# using the glmmadmb package to create a zero inflation mixed linear poisson plot 

footds = glmmadmb(formula = foot_flags ~ treatment * hour + (1 | fterrarium), data = footds, family = "truncnbinom", zeroInflation = TRUE) 
summary(footds)
```

```{r}
# verifying model
hist(exp2$foot_flags)
```

```{r}
poisson.test(sum(exp2$foot_flags), length(exp2$foot_flags))
```


# Making a model for the foot flashing behavior 

```{r}
#selected for only the foot flashing behavior. short for "flashes data set"
flashds <- exp2 %>%
  select(treatment, foot_flashes, hour, terrarium, video_name)
```

```{r}
# made terrarium variable a factor
flashds <- transform(flashds,fterrarium=factor(terrarium))
```

```{r}
# using the glmmadmb package to create a zero inflation mixed linear poisson plot 
# flashds = glmmadmb(formula = foot_flashes ~ treatment * hour + (1 | fterrarium), data = flashds, family = "truncnbinom", zeroInflation = TRUE) 
# summary(flashds)
```

```{r}
# verifying model
hist(exp2$foot_flashes)
```

```{r}
poisson.test(sum(exp2$foot_flashes), length(exp2$flashes))
```


# Filtering Hours for Experiment 1
```{r}
filterhour1 <-subset(exp2, hour == 1)
summary(filterhour1)

# made terrarium variable a factor
filterhour1 <- transform(filterhour1,fterrarium =factor(terrarium))
```
```{r}
# Code for hour 2:
filterhour2 <-subset(exp2, hour == 2)
#summary(filterhour2)

# made terrarium variable a factor
filterhour2 <- transform(filterhour2,fterrarium =factor(terrarium))

#Code for hour 3: 

filterhour3 <-subset(exp2, hour == 3)
#summary(filterhour3)

# made terrarium variable a factor
filterhour3 <- transform(filterhour3,fterrarium =factor(terrarium))
```

# Run T test for each hour with random variable for experiment 2

## Hour 1
```{r}
# for movements
hour1_movements = glmmadmb(formula = total_major_movements ~ treatment + (1 | fterrarium),data = filterhour1, family ="truncnbinom" , zeroInflation = TRUE)
summary(hour1)
```

```{r}
# for touches
hour1_touches = glmmadmb(formula = touches ~ treatment + (1 | fterrarium),data = filterhour1, family ="poisson" , zeroInflation = TRUE)
summary(hour1_touches)
```

```{r}
# for foot_flags
hour1_foot_flags = glmmadmb(formula = foot_flags~ treatment + (1 | fterrarium),data = filterhour1, family ="truncnbinom" , zeroInflation = TRUE)
summary(hour1_foot_flags)
```

```{r}
# for foot flashes
hour1_flash = glmmadmb(formula = foot_flashes ~ treatment + (1 | fterrarium),data = filterhour1, family ="truncnbinom" , zeroInflation = TRUE)
summary(hour1_flash)
```

## Hour 2
```{r}
# for movements
# hour2_movements = glmmadmb(formula = total_major_movements ~ treatment + (1 | fterrarium),data = filterhour2, family ="truncnbinom" , zeroInflation = TRUE)
# summary(hour2_movements)
```


```{r}
# for touches
hour2_touches = glmmadmb(formula = touches ~ treatment + (1 | fterrarium),data = filterhour2, family ="poisson" , zeroInflation = TRUE)
summary(hour2_touches)
```

```{r}
# for foot_flags
hour2_foot_flags = glmmadmb(formula = foot_flags~ treatment + (1 | fterrarium),data = filterhour2, family ="truncnbinom" , zeroInflation = TRUE)
summary(hour2_foot_flags)
```

```{r}
# for foot flashes
hour2_flash = glmmadmb(formula = foot_flashes ~ treatment + (1 | fterrarium),data = filterhour1, family ="truncnbinom" , zeroInflation = TRUE)
summary(hour2_flash)
```

## Hour 3

```{r}
# for movements
hour3_movements = glmmadmb(formula = total_major_movements ~ treatment + (1 | fterrarium),data = filterhour3, family ="truncnbinom" , zeroInflation = TRUE)
summary(hour3_movements)
```

```{r}
# for touches
hour3_touches = glmmadmb(formula = touches ~ treatment + (1 | fterrarium),data = filterhour3, family ="poisson" , zeroInflation = TRUE)
summary(hour3_touches)
```

```{r}
# for foot_flags
hour3_foot_flags = glmmadmb(formula = foot_flags~ treatment + (1 | fterrarium),data = filterhour3, family ="truncnbinom" , zeroInflation = TRUE)
summary(hour3_foot_flags)
```

```{r}
# for foot flashes
hour3_flash = glmmadmb(formula = foot_flashes ~ treatment + (1 | fterrarium),data = filterhour3, family ="truncnbinom" , zeroInflation = TRUE)
summary(hour3_flash)
```

# Running Comparison without random variable for experiment 2:

## Hour 1: 
```{r}
# for total major movements
newhour1_movements = glmmadmb(formula = total_major_movements ~ treatment,data = filterhour1, family = "poisson", zeroInflation = TRUE)
summary(newhour1_movements)
```


```{r}
# for touches 
newhour1_touches= glmmadmb(formula = touches ~ treatment,data = filterhour1, family = "poisson", zeroInflation = TRUE)
summary(newhour1_touches)
```

```{r}
# for foot flags
newhour1_foot = glmmadmb(formula = foot_flags~ treatment,data = filterhour1, family = "poisson", zeroInflation = TRUE)
summary(newhour1_foot)
```

```{r}
# for foot flashes
newhour1_flashes = glmmadmb(formula = foot_flashes ~ treatment,data = filterhour1, family = "poisson", zeroInflation = TRUE)
summary(newhour1_flashes)
```



## Hour 2: 
```{r}
# for total major movements
newhour2_movements = glmmadmb(formula = total_major_movements ~ treatment,data = filterhour2, family = "poisson", zeroInflation = TRUE)
summary(newhour2_movements)
```

```{r}
# for touches 
newhour2_touches= glmmadmb(formula = touches ~ treatment,data = filterhour2, family = "poisson", zeroInflation = TRUE)
summary(newhour2_touches)
```

```{r}
# for foot flags
newhour2_foot = glmmadmb(formula = foot_flags~ treatment,data = filterhour2, family = "poisson", zeroInflation = TRUE)
summary(newhour2_foot)
```

```{r}
# for foot flashes
newhour2_flashes = glmmadmb(formula = foot_flashes ~ treatment,data = filterhour2, family = "poisson", zeroInflation = TRUE)
summary(newhour2_flashes)
```

## Hour 3: 
```{r}
# for total major movements
newhour3_movements = glmmadmb(formula = total_major_movements ~ treatment,data = filterhour3, family = "poisson", zeroInflation = TRUE)
summary(newhour3_movements)
```


```{r}
# for touches 
newhour3_touches= glmmadmb(formula = touches ~ treatment,data = filterhour3, family = "poisson", zeroInflation = TRUE)
summary(newhour3_touches)
```

```{r}
# for foot flags
newhour3_foot = glmmadmb(formula = foot_flags~ treatment,data = filterhour3, family = "poisson", zeroInflation = TRUE)
summary(newhour3_foot)
```

```{r}
# for foot flashes
newhour3_flashes = glmmadmb(formula = foot_flashes ~ treatment,data = filterhour3, family = "poisson", zeroInflation = TRUE)
summary(newhour3_flashes)
```

# With the pscl package

```{r}
# What % of the data is zeroes? Grouped by hour, then by treatment
##tally(~foot_flags==0 | treatment, groups=hour, format="percent", data=exp2)
```

```{r}
# Create a model called “zero1” which uses the zero-inflated regression model
# Dependent variable is count data = Behavior
# Predictor variable is “Treatment” coded as “AVT" or “Saline”

zero1_m <- zeroinfl(total_major_movements ~ treatment, data = exp2)
summary(zero1)
```
```{r}
# zero-inflated model with no interaction for behavior = touches
zero1_t <- zeroinfl(touches ~ treatment, data = exp2)
summary(zero1_t)
```

```{r}
# zero-inflated model with no interaction for behavior = foot flags
zero1_foot <- zeroinfl(foot_flags ~ treatment, data = exp2)
summary(zero1_foot)
```

```{r}
# zero-inflated model with no interaction for behavior = foot flashes
zero1_flash <- zeroinfl(foot_flashes ~ treatment, data = exp2)
summary(zero1_flash)
```

```{r}
# Full model with interaction term, behavior = movements
m2 <- zeroinfl(total_major_movements ~ treatment + treatment*hour, data = exp2)
summary(m2)
```

```{r}
# Full model with interaction term, behavior = touches 
t2 <- zeroinfl(touches~ treatment + treatment*hour, data = exp2)
summary(t2)
```

```{r}
# Full model with interaction term, behavior = movements
foot2 <- zeroinfl(foot_flags ~ treatment + treatment*hour, data = exp2)
summary(foot2)
```

```{r}
# Full model with interaction term, behavior = movements
flash2 <- zeroinfl(foot_flashes ~ treatment + treatment*hour, data = exp2)
summary(flash2)
```


# Focusing on Hour 2, using the pscl model 

```{r}
#for movements: get an hour, says "invalid dependent variable, minimum count is not zero"
# zeroinflhour2m <- zeroinfl(total_major_movements ~ treatment, data = filterhour2)
# summary(zeroinflhour2)
```
```{r}
# for touches behavior 
zeroinflhour2t <- zeroinfl(touches~ treatment, data = filterhour2)
summary(zeroinflhour2t)
```

```{r}
# for foot flags
# zeroinflhour2foot <- zeroinfl(foot_flags ~ treatment, data = filterhour2)
# summary(zeroinflhour2foot)
```


```{r}
# for foot flashes
# zeroinflhour2flash <- zeroinfl(foot_flashes ~ treatment, data = filterhour2)
# summary(zeroinflhour2flash)
```
